Generalized Plackett-Luce Likelihoods

### 2022-08-22 inst/jss_hyper3.Rnw
spaces after commas

<<>>=
LL <- hyper3()
LL[c(a=1)] <- 1
LL[c(a=3,b=2,c=1)] <- -1
LL
@

@@

<<>>=
loglik(c(a=0.01,b=0.01,c=0.98),LL)
loglik(c(a=0.9,b=0.05,c=0.05),LL)
@

@@

<<>>=
(H <- ordervec2supp3(c("a","c","b","a","a","b")))
@

@@

<<maxpaba,cache=TRUE>>=
maxp(ordervec2supp3(c("a","b","a")))
@

@@

\begin{figure}[htbp]
  \begin{center}
<<figabbabb,echo=FALSE>>=
f_aab <- function(a){a^2/(1+a)}
f_aba <- function(a){a*(1-a)/(1+a)^2}
f_baa <- function(a){(1-a)/(1+a)}

p <- function(f,...){points(a,f(a)/max(f(a)),...)}
plot(0:1,0:1,xlab="p(a)",ylab="Likelihood",type="n")
p(f_aab,type="l",col="black")
p(f_aba,type="l",col="red")
p(f_baa,type="l",col="blue")
legend(x=0.85,y=0.7,col=c("black","red","blue"),lty=1,pch=NA,legend=c("AAB","ABA","BAA"))
abline(h=exp(-2),lty=2)
@
\caption{Likelihood \label{ABAetc_likelihood} functions for
observations $a\succ a\succ b$, $a\succ b\succ a$, $b\succ a\succ a$}
\end{center}
\end{figure}

@@

<<cache=TRUE>>=
maxp(ordervec2supp3(c("a","b"),nonfinishers=c("a","b")))
@

@@

<<hyper3osdef,cache=TRUE>>=
names(x) <- rep("x",length(x))
names(y) <- rep("y",length(y))
(os <- names(sort(c(x,y))))
@

@@

\begin{figure}[htbp]
  \begin{center}
<<plotwilcoxlike,echo=FALSE>>=
a <- seq(from=0.02,to=0.8,len=40)
L <- sapply(a,function(p){loglik(p,Hxy)})
plot(a,L-max(L),type='b')
abline(h=c(0,-2))
abline(v=c(0.24))
abline(v=c(0.5),lty=2)
@
\caption{A\label{wilcox_likelihood} support for the Bradley-Terry
strength $p_a$ of permeability at term.  The evaluate of 0.24 is
shown; and the two-units-of support credible interval, which does not
exclude $H_0\colon p_a=0.5$ (dotted line), is also shown}
\end{center}
\end{figure}

@@

<<label=converttosupp3,cache=TRUE>>=
javelin_vector <- attemptstable2supp3(javelin_table,decreasing=TRUE,give.supp=FALSE)
javelin_vector
@

@@

\begin{figure}[htbp]
  \begin{center}
<<testthejav,cache=TRUE>>=
(mj <- maxp(javelin))
dotchart(mj,pch=16)
@
\caption{Maximum likelihood estimate for javelin throwers' Bradley-Terry strengths}
\end{center}
\end{figure}

@@

<<havealook,echo=FALSE,cache=TRUE>>=
f <- function(s){
  jj <- specificp.test(javelin,"Vesely",s,n=2)
  p <- jj$null_estimate
 return(c(log(p[6]/p[7]),jj$null_support))
}
Ves <- seq(from=0.0199,to=0.33,len=16)
M <- sapply(Ves,f)
M[2,] <- M[2,]-max(M[2,])
rownames(M) <- c("logcontrast","support")
@

\begin{figure}[htbp]
\begin{center}
<<plottheloglikcont,echo=FALSE>>=
colnames(M) <-  as.character(Ves)
plot(t(M),type="b")
abline(h=c(0,-2))
abline(v=0,lty=2)
abline(v=log(0.32062833/0.11402735)) # these from mp
@
\caption{Profile \label{profliklogcont} likelihood for log-contrast
$\log\left(p_{\mbox{\tiny Vad}}/ p_{\mbox{\tiny Ves}}\right)$.  Null
of $p_{\mbox{\tiny Vad}}=p_{\mbox{\tiny Ves}}$ indicated with a dotted
line, and two-units-of-support limit indicated with horizontal lines;
thus the null is not rejected}
\end{center}
\end{figure}

@@

<<maxpconstructor2021, cache=TRUE>>=
const2020 <- ordertable2supp3(constructor_2020_table)
const2021 <- ordertable2supp3(constructor_2021_table)
maxp(const2020,n=1)
maxp(const2021,n=1)
@

@@

<<definecombinedlikelihoodfunction,cache=TRUE>>=
H <- psubs(constructor_2020,"Merc","Merc2020") + psubs(constructor_2021,"Merc","Merc2021")
@

@@

<<usesamep,cache=TRUE>>=
samep.test(H,c("Merc2020","Merc2021"))
@

>>>>

<<>>=
LL <- hyper3()
LL[c(a=1)] <- 1
LL[c(a=3, b=2, c=1)] <- -1
LL
@

@@

<<>>=
loglik(c(a=0.01, b=0.01, c=0.98),LL)
loglik(c(a=0.90, b=0.05, c=0.05),LL)
@

@@

<<>>=
(H <- ordervec2supp3(c("a", "c", "b", "a", "a", "b")))
@

@@

<<maxpaba,cache=TRUE>>=
maxp(ordervec2supp3(c("a", "b", "a")))
@

@@

\begin{figure}[htbp]
  \begin{center}
<<figabbabb,echo=FALSE>>=
f_aab <- function(a){a^2 / (1 + a)}
f_aba <- function(a){a * (1 - a) / (1 + a)^2}
f_baa <- function(a){(1 - a) / (1 + a)}

p <- function(f, ...){points(a, f(a) / max(f(a)), ...)}
plot(0:1, 0:1, xlab="p(a)", ylab="Likelihood", type="n")
p(f_aab, type="l", col="black")
p(f_aba, type="l", col="red")
p(f_baa, type="l", col="blue")
legend(x=0.85, y=0.7, col=c("black", "red", "blue"), lty=1, pch=NA, legend=c("AAB", "ABA", "BAA"))
abline(h=exp(-2), lty=2)
@
\caption{Likelihood \label{ABAetc_likelihood} functions for
observations $a\succ a\succ b$, $a\succ b\succ a$, $b\succ a\succ a$}
\end{center}
\end{figure}

@@

<<cache=TRUE>>=
maxp(ordervec2supp3(c("a", "b"), nonfinishers=c("a", "b")))
@

@@

<<hyper3osdef,cache=TRUE>>=
names(x) <- rep("x", length(x))
names(y) <- rep("y", length(y))
(os <- names(sort(c(x, y))))
@

@@

\begin{figure}[htbp]
  \begin{center}
<<plotwilcoxlike,echo=FALSE>>=
a <- seq(from=0.02, to=0.8, len=40)
L <- sapply(a, function(p){loglik(p, Hxy)})
plot(a, L - max(L),type='b')
abline(h=c(0, -2))
abline(v=c(0.24))
abline(v=c(0.5), lty=2)
@
\caption{A\label{wilcox_likelihood} support for the Bradley-Terry
strength $p_a$ of permeability at term.  The evaluate of 0.24 is
shown; and the two-units-of support credible interval, which does not
exclude $H_0\colon p_a=0.5$ (dotted line), is also shown}
\end{center}
\end{figure}

@@

<<label=converttosupp3,cache=TRUE>>=
javelin_vector <- attemptstable2supp3(javelin_table, decreasing=TRUE, give.supp=FALSE)
javelin_vector
@

@@

\begin{figure}[htbp]
  \begin{center}
<<testthejav,cache=TRUE>>=
(mj <- maxp(javelin))
dotchart(mj, pch=16)
@
\caption{Maximum likelihood estimate for javelin throwers' Bradley-Terry strengths}
\end{center}
\end{figure}

@@

<<havealook,echo=FALSE,cache=TRUE>>=
f <- function(s){
  jj <- specificp.test(javelin, "Vesely", s, n=2)
  p <- jj$null_estimate
 return(c(log(p[6] / p[7]), jj$null_support))
}
Ves <- seq(from=0.0199, to=0.33, len=16)
M <- sapply(Ves, f)
M[2,] <- M[2,] - max(M[2,])
rownames(M) <- c("logcontrast", "support")
@

\begin{figure}[htbp]
\begin{center}
<<plottheloglikcont,echo=FALSE>>=
colnames(M) <-  as.character(Ves)
plot(t(M), type="b")
abline(h=c(0, -2))
abline(v=0, lty=2)
abline(v=log(0.32062833/0.11402735)) # these from mp
@
\caption{Profile \label{profliklogcont} likelihood for log-contrast
$\log\left(p_{\mbox{\tiny Vad}}/ p_{\mbox{\tiny Ves}}\right)$.  Null
of $p_{\mbox{\tiny Vad}}=p_{\mbox{\tiny Ves}}$ indicated with a dotted
line, and two-units-of-support limit indicated with horizontal lines;
thus the null is not rejected}
\end{center}
\end{figure}

@@

<<maxpconstructor2021, cache=TRUE>>=
const2020 <- ordertable2supp3(constructor_2020_table)
const2021 <- ordertable2supp3(constructor_2021_table)
maxp(const2020, n=1)
maxp(const2021, n=1)
@

@@

<<definecombinedlikelihoodfunction,cache=TRUE>>=
H <- psubs(constructor_2020, "Merc", "Merc2020") + psubs(constructor_2021, "Merc", "Merc2021")
@

@@

<<usesamep,cache=TRUE>>=
samep.test(H, c("Merc2020", "Merc2021"))
@
